Prediction of *PIK3CA* mutations from cancer gene expression data

### 2020-07-22 plosOne/plosOne.Rmd
submition

abstract: | 
  Breast cancers with *PIK3CA* mutations can be treated with *PIK3CA* inhibitors in hormone receptor-positive HER2 negative subtypes. We applied a supervised elastic net penalized logistic regression model to predict *PIK3CA* mutations from gene expression data. This regression approach was applied to predict modeling using the TCGA pan-cancer dataset. Approximately 10,000 cases were available for *PIK3CA* mutation and mRNA expression data. In 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (Ridge regression) showed the best performance, in terms of area under the receiver operating characteristic (AUROC). The final model was developed with selected hyper-parameters using the entire training set. The training set AUROC was 0.93, and the test set AUROC was 0.84. The area under the precision-recall (AUPR) of the training set was 0.66, and the test set AUPR was 0.39. Cancer types were the most important predictors. Both *insulin like growth factor 1 Receptor* (*IGF1R*) and the *phosphatase and tensin homolog* (*PTEN*) were the most significant genes in gene expression predictors. Our study suggests that predicting genomic alterations using gene expression data is possible, with good outcomes.

@@

For 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (Ridge regression) showed the best performance in terms of AUROC. The final model was trained with the selected hyper-parameters with the entire training set. The training set AUROC was 0.93 and the test set AUROC was 0.84. The AUPR of the training set was 0.66 and the test set AUPR was 0.39 (Figure 2A).

>>>>

abstract: | 
  Breast cancers with *PIK3CA* mutations can be treated with *PIK3CA* inhibitors in hormone receptor-positive HER2 negative subtypes. We applied a supervised elastic net penalized logistic regression model to predict *PIK3CA* mutations from gene expression data. This regression approach was applied to predict modeling using the TCGA pan-cancer dataset. Approximately 10,000 cases were available for *PIK3CA* mutation and mRNA expression data. In 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (ridge regression) showed the best performance, in terms of area under the receiver operating characteristic (AUROC). The final model was developed with selected hyper-parameters using the entire training set. The training set AUROC was 0.93, and the test set AUROC was 0.84. The area under the precision-recall (AUPR) of the training set was 0.66, and the test set AUPR was 0.39. Cancer types were the most important predictors. Both *insulin like growth factor 1 receptor* (*IGF1R*) and the *phosphatase and tensin homolog* (*PTEN*) were the most significant genes in gene expression predictors. Our study suggests that predicting genomic alterations using gene expression data is possible, with good outcomes.

@@

For 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (ridge regression) showed the best performance in terms of AUROC. The final model was trained with the selected hyper-parameters with the entire training set. The training set AUROC was 0.93 and the test set AUROC was 0.84. The AUPR of the training set was 0.66 and the test set AUPR was 0.39 (Figure 2A).

### 2020-07-22 plosOne/plosOne.tex
submition

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Prediction of \emph{PIK3CA} mutations from cancer gene expression data} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Jun Kang\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}},
Ahwon Lee\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}},
Youn Soo Lee\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}}\textsuperscript{*}\\
\bigskip
\textbf{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}}Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
Medicine, The Catholic University of Korea, Seoul, South Korea\\
\bigskip
* Corresponding author: lys9908@catholic.ac.kr\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Breast cancers with \emph{PIK3CA} mutations can be treated with
\emph{PIK3CA} inhibitors in hormone receptor-positive HER2 negative
subtypes. We applied a supervised elastic net penalized logistic
regression model to predict \emph{PIK3CA} mutations from gene expression
data. This regression approach was applied to predict modeling using the
TCGA pan-cancer dataset. Approximately 10,000 cases were available for
\emph{PIK3CA} mutation and mRNA expression data. In 10-fold
cross-validation, the model with \(\lambda\) = 0.01 and \(\alpha\) = 1.0
(Ridge regression) showed the best performance, in terms of area under
the receiver operating characteristic (AUROC). The final model was
developed with selected hyper-parameters using the entire training set.
The training set AUROC was 0.93, and the test set AUROC was 0.84. The
area under the precision-recall (AUPR) of the training set was 0.66, and
the test set AUPR was 0.39. Cancer types were the most important
predictors. Both \emph{insulin like growth factor 1 Receptor}
(\emph{IGF1R}) and the \emph{phosphatase and tensin homolog}
(\emph{PTEN}) were the most significant genes in gene expression
predictors. Our study suggests that predicting genomic alterations using
gene expression data is possible, with good outcomes.

@@

For 10-fold cross-validation, the model with \(\lambda\) = 0.01 and
\(\alpha\) = 1.0 (Ridge regression) showed the best performance in terms
of AUROC. The final model was trained with the selected hyper-parameters
with the entire training set. The training set AUROC was 0.93 and the
test set AUROC was 0.84. The AUPR of the training set was 0.66 and the
test set AUPR was 0.39 (Figure 2A).

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Prediction of \emph{PIK3CA} mutations from cancer gene expression data} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Jun Kang\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}},
Ahwon Lee\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}},
Youn Soo Lee\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}}\textsuperscript{*}\\
\bigskip
\textbf{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}}Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
Medicine, The Catholic University of Korea, Seoul, South Korea\\
\bigskip
* Corresponding author: lys9908@catholic.ac.kr\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Breast cancers with \emph{PIK3CA} mutations can be treated with
\emph{PIK3CA} inhibitors in hormone receptor-positive HER2 negative
subtypes. We applied a supervised elastic net penalized logistic
regression model to predict \emph{PIK3CA} mutations from gene expression
data. This regression approach was applied to predict modeling using the
TCGA pan-cancer dataset. Approximately 10,000 cases were available for
\emph{PIK3CA} mutation and mRNA expression data. In 10-fold
cross-validation, the model with \(\lambda\) = 0.01 and \(\alpha\) = 1.0
(ridge regression) showed the best performance, in terms of area under
the receiver operating characteristic (AUROC). The final model was
developed with selected hyper-parameters using the entire training set.
The training set AUROC was 0.93, and the test set AUROC was 0.84. The
area under the precision-recall (AUPR) of the training set was 0.66, and
the test set AUPR was 0.39. Cancer types were the most important
predictors. Both \emph{insulin like growth factor 1 receptor}
(\emph{IGF1R}) and the \emph{phosphatase and tensin homolog}
(\emph{PTEN}) were the most significant genes in gene expression
predictors. Our study suggests that predicting genomic alterations using
gene expression data is possible, with good outcomes.

@@

For 10-fold cross-validation, the model with \(\lambda\) = 0.01 and
\(\alpha\) = 1.0 (ridge regression) showed the best performance in terms
of AUROC. The final model was trained with the selected hyper-parameters
with the entire training set. The training set AUROC was 0.93 and the
test set AUROC was 0.84. The AUPR of the training set was 0.66 and the
test set AUPR was 0.39 (Figure 2A).
